Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto

### 2024-03-16 paper/paper.qmd
finished final rendering

---
title: "Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

-   On average, for every increase in calls coded by the helpline, there
    is a 0.0003 unit decrease in homeless death in Toronto. There is an
    overall reduction, albeit very small, in homeless death as the
    number of total coded calls increase. This implies that the helpline
    need to take at minimum 3333 calls per month to save 1 homeless life
    in Toronto. On average, the helpline takes 14828.59 calls per month,
    meaning that 4.45 lives are saved on average.
-   4.45 is not a significant number in both a practical and statistical
    sense. To improve the return on total number of calls coded, we
    attempted to identify promising call codes that contribute to the
    net positive reduction in homeless death. Out of the 13 different
    call codes, *Referral to Shelter* and *Information on Homeless
    Prevention* are best suited as individual efforts of measurement for
    positive impacts because they provide a count of the number of
    callers provided with directions or advice. For every caller
    provided with information on homeless prevention information, there
    is an average decrease of 0.0048 death count in Toronto. On the
    other hand, for every homeless caller referred to shelter, there is
    a 0.0059 unit of average **increase** in homeless death in Toronto.
    This trend is very counter-intuitive and discouraging at first
    glance. In the case where the correlation is more significant, we
    would want to further analyze factor related to referrals to
    shelter. Accuracy on the current status of shelters, level of detail
    on instructed, and misjudgment of shelter eligibility are possible
    factors of examination. However, since the effect of this call code
    is statistically insignificant, it is also likely to be just a fluke
    of observation.

@@

In @fig-total-coded-model-ppcheckandposteriorvsprior-1 we implement a
posterior predictive check. This shows a relatively close match between
the regression outcome from our data and the simulation created by the
posterior distribution. This suggests that our model is a relatively
good fit for our data.

In @fig-total-coded-model-ppcheckandposteriorvsprior-2 we compare the
posterior with the prior. We can see that the estimation parameters
shifts minimally after taking data into account. This suggests that good
prior parameters were set.

@@

In @fig-referred-informed-model-ppcheckandposteriorvsprior-1 we
implement a posterior predictive check. This shows a relatively close
match between the regression outcome from our data and the simulation
created by the posterior distribution. This suggests that our model is a
relatively good fit for our data.

In @fig-referred-informed-model-ppcheckandposteriorvsprior-2 we compare
the posterior with the prior. We can see that the estimation parameters
shifts minimally after taking data into account. This suggests that good
prior parameters were set.

@@

@fig-total-coded-model-stanareyouokay-1 is a trace plot. Trace plots are
often used to diagnose abnormalities of models. In
@fig-total-coded-model-stanareyouokay-1, lines are oscillating
vertically and maintains an overall horizontal trend. This suggests that
there is nothing out of the ordinary for our model.

@fig-total-coded-model-stanareyouokay-2 is a Rhat plot. Rhat plot is a
valuable tool for assessing convergence in Bayesian regression models,
helping ensure the reliability and validity of the model's inference
results. In @fig-total-coded-model-stanareyouokay-2, all data points are
close to 1. This suggest that there is nothing out of the ordinary for
our model.

@@

@fig-referred-informed-model-stanareyouokay-1 is a trace plot. Trace
plots are often used to diagnose abnormalities of models. In
@fig-referred-informed-model-stanareyouokay-1, lines are oscillating
vertically and maintains an overall horizontal trend. This suggests that
there is nothing out of the ordinary for our model.

@fig-referred-informed-model-stanareyouokay-2 is a Rhat plot. Rhat plot
is a valuable tool for assessing convergence in Bayesian regression
models, helping ensure the reliability and validity of the model's
inference results. In @fig-referred-informed-model-stanareyouokay-2, all
data points are close to 1. This suggest that there is nothing out of
the ordinary for our model.

>>>>

---
title: "Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

On average, for every increase in calls coded by the helpline, there is a 0.0003 unit decrease in homeless death in Toronto. There is an overall reduction, albeit very small, in homeless death as the number of total coded calls increase. This implies that the helpline need to take at minimum 3333 calls per month to save 1 homeless life in Toronto. On average, the helpline takes 14828.59 calls per month, meaning that 4.45 lives are saved on average.

4.45 is not a significant number in both a practical and statistical sense. To improve the return on total number of calls coded, we attempted to identify promising call codes that contribute to the net positive reduction in homeless death. Out of the 13 different call codes, *Referral to Shelter* and *Information on Homeless Prevention* are best suited as individual efforts of measurement for positive impacts because they provide a count of the number of callers provided with directions or advice. For every caller provided with information on homeless prevention information, there is an average decrease of 0.0048 death count in Toronto. On the other hand, for every homeless caller referred to shelter, there is a 0.0059 unit of average **increase** in homeless death in Toronto. This trend is very counter-intuitive and discouraging at first glance. In the case where the correlation is more significant, we would want to further analyze factor related to referrals to shelter. Accuracy on the current status of shelters, level of detail on instructed, and misjudgment of shelter eligibility are possible factors of examination. However, since the effect of this call code is statistically insignificant, it is also likely to be just a fluke of observation.

@@

In @fig-total-coded-model-ppcheckandposteriorvsprior-1 and @fig-referred-informed-model-ppcheckandposteriorvsprior-1, we implement a posterior predictive check. This shows a relatively close match between the regression outcome from our data and the simulation created by the posterior distribution. This suggests that our model is a relatively good fit for our data.

In @fig-total-coded-model-ppcheckandposteriorvsprior-2 and @fig-referred-informed-model-ppcheckandposteriorvsprior-2, we compare the posterior with the prior. We can see that the estimation parameters shifts minimally after taking data into account. This suggests that good prior parameters were set.

@@

@@

@fig-total-coded-model-stanareyouokay-1 and @fig-total-coded-model-stanareyouokay-2 are trace plots. Trace plots are often used to diagnose abnormalities of models. For both plots, lines are oscillating vertically and maintains an overall horizontal trend. This suggests that there is nothing out of the ordinary for our model.

@fig-total-coded-model-stanareyouokay-2 and @fig-referred-informed-model-stanareyouokay-2 are Rhat plots. Rhat plot is a valuable tool for assessing convergence in Bayesian regression models, helping ensure the reliability and validity of the model's inference results. Both both plots, all data points are close to 1. This suggest that there is nothing out of the ordinary for our model.

@@
